---
title: "Medical Insurance Project"
author: "By Jesse"
output: html_document
---

##### In this I project I explore health insurance data, attempt to identify which factors are most important in predicting medical expenses, and perform linear regression analysis.

##### Below are the factors provided in the Medical Cost data set:

* age: age of primary beneficiary
* sex: insurance contractor gender, female, male
* bmi: Body mass index, providing an understanding of body, weights that are relatively high or low relative to height, objective index of body weight (kg / m ^ 2) using the ratio of height to weight, ideally 18.5 to 24.9
* children: Number of children covered by health insurance / Number of dependents
* smoker: Smoking
* region: the beneficiary's residential area in the US, northeast, southeast, southwest, northwest.
* charges: Individual medical costs billed by health insurance

<br>

Let's get started!

Loading in packages, removing objects from the environment and setting the working directory.

```{r packages and wd, message=FALSE, warning=FALSE}
if (!require("pacman")) install.packages("pacman"); library(pacman)
pacman::p_load(tidyverse, caTools, ggthemes, janitor, corrplot, GGally, psych)

rm(list=ls())
setwd("/Users/Jesse/R/MedInsuranceProject/")
```

Reading in the data.

```{r reading in the data, message=FALSE, warning=FALSE}
df <- read_csv("/Users/Jesse/R/MedInsuranceProject/insurance.csv")
```

Let's take a took at the data:

```{r checking out the data}
head(df)
str(df)
```

Let's change the character columns to factors:

```{r factors}
df$sex <- as.factor(df$sex)
df$smoker <- as.factor(df$smoker)
df$region <- as.factor(df$region)
```

Finally, let's check for any missing values:

```{r missing}
any(is.na(df))
```

No missing data! Let's move on to some exploratory data analysis.

<br>

#### EDA

```{r EDA summary}
summary(df)
```

Let's take a look at a histogram of the insurance charges:

```{r EDA histogram 1}
df %>% ggplot(aes(charges)) + geom_histogram(aes(y=stat(density)), fill = "light blue", color = "white", bins = 30) + 
    ggtitle("Distribution of Charges") + theme_bw() + theme(plot.title = element_text(hjust = .5)) +
    geom_density(col="dark blue")
```

<br>

This distribution is right-skewed. Let's make the distribution closer to normal:

```{r EDA histogram 2}
df %>% ggplot(aes(log10(charges))) + geom_histogram(aes(y=stat(density)), fill = "light blue", color = "white", bins = 30) + 
    ggtitle("Distribution of Charges") + theme_bw() + theme(plot.title = element_text(hjust = .5)) + 
    geom_density(col = "dark blue")
```

<br>

Now let's look at charges by region:

```{r EDA region}
describeBy(df$charges,df$region)

df %>% ggplot(aes(region, charges)) + geom_boxplot(fill=c(3,4,6,7)) + 
    theme_bw() + ggtitle("Medical Insurance Charges by Region")
```

<br>

It looks like charges don't differ much across regions, but the highest medical expenses are in the Southeast and the lowest are in the Southwest.

Let's check out charges by smoker status:

```{r EDA smoker}
describeBy(df$charges,df$smoker)

df %>% ggplot(aes(smoker, charges)) + geom_boxplot(fill=c(3,2)) + 
    theme_bw() + ggtitle("Medical Insurance Charges by Smoker Status")
```

<br>

As you would expect, smokers spend much more than non-smokers. Smokers spend about 4x more compared to non-smokers when it comes to medical insurance expenses.

Next, let's look at sex:

```{r EDA sex}
describeBy(df$charges,df$sex)

df %>% ggplot(aes(sex, charges)) + geom_boxplot(fill=c(5,7)) + 
    theme_bw() + ggtitle("Medical Insurance Charges by Sex")
```

<br>

It appears sex doesn't seem to affect medical expenses very much.

Moving on to number of children:

```{r EDA children}
describeBy(df$charges,df$children)

df %>% ggplot(aes(as.factor(children), charges)) + geom_boxplot(fill=c(2:7)) + 
    theme_bw() + ggtitle("Medical Insurance Charges by Number of Children")
```

<br>

Interestingly, on average, people with 5 children spend less on medical expense than all other groups.

Let's check out obesity now. First we need to create a new column for obese, which is greater than or equal to a BMI of 30.

```{r EDA obesity 1}
df <- df %>%
    mutate(obese = as.factor(if_else(bmi >= 30, "yes", "no")))

describeBy(df$charges,df$obese)

df %>% ggplot(aes(obese, charges)) + geom_boxplot(fill=c(6,4)) + 
    theme_bw() + ggtitle("Medical Charges by Obesity status (BMI 30+)")
```

<br>

As we can see, the average medical expenses for an obese individual is about $5000 more than non-obese people,
while the median expenses between the group is similar.

Lastly, let's look at age, but also with smoker status.

```{r EDA age}
df %>% ggplot(aes(age, charges)) + geom_point(aes(color = smoker)) +
    theme_bw() + ggtitle("Medical Charges by Age and Smoker Status") + scale_color_manual(values=c("red", "blue"))
```

<br>

As expected, as people get older they tend to have higher medical expenses, but adding smoker status you can see the relationship between smoking and higher medical expenses.

Alright, let's move on to looking at correlation between variables.

<br>

#### Correlation

```{r EDA cor 1}
num_cols <- sapply(df,is.numeric) # using only numeric features
ggpairs(df[,num_cols]) # matrix of plots
pairs.panels(df[,num_cols]) # matrix of plots using pairs.panels
```

We can see that age and charges have the highest correlation among the numeric variables. It's also interesting 
to note that none of our numeric values are highly correlated with each other, so multicollinearity won’t be a problem.

Now let's look at the correlation matrix for numeric features a different way.

```{r EDA cor 2}
cor_mx <- cor(df[,num_cols]) # correlation matrix
cor_mx
corrplot(cor_mx, method = "color", title = "Correlation Matrix for Numeric Features", 
         mar=c(0,0,1,0))
```

<br>

Let's move on and look at correlation between all features. But first we need to create dummy variables for sex and smoker.

```{r EDA cor 3}
df_dummy <- df %>%
    mutate(sex_dummy = if_else(sex == "male", 1, 0)) %>%
    mutate(smoker_dummy = if_else(smoker == "yes", 1, 0))

df_dummy <- df_dummy %>% 
    select(1,3,4,7,9,10)

# correlation all features
ggpairs(df_dummy)
pairs.panels(df_dummy)

dummy_cor_mx <- cor(df_dummy)
corrplot(dummy_cor_mx, method = "color", title = "Correlation Matrix for All Features", 
         mar=c(0,0,1,0))
```

<br>

We can see that smoking and medical expense are highly correlated.

Let's move on to building a model using all of the original features, except using obese instead of BMI because BMI is essentially the same thing as obese.

<br>

#### Building a model

```{r model 1}
set.seed(214)
sample <- sample.split(df$charges, SplitRatio = 0.70)

# Training Data
train = subset(df, sample == TRUE)

# Testing Data
test = subset(df, sample == FALSE)

med_model <- lm(charges ~ .-bmi, data = train)
summary(med_model)
plot(med_model)
```

<br>

Just using the original variables we get a fairly good R-squared of 0.7304, which implies that about 73% of the variation in charges can be explained by the independent variables. Also, all variables except for region and sex, are statistically significant predictors of medical expense (p < .05).

<br>

Let's attempt to improve the model:

First, let's remove region and sex, as they were not statistically significant.

```{r model 2}
med_model2 <- lm(charges ~ . -bmi -sex -region, data = train)
summary(med_model2)
plot(med_model2)
```

<br>

This model is about the same, but slightly worse, with an R-squared of 0.729, still implying about 73% of the variation in charges can be explained by the independent variables. Let's try a model with an interaction between obesity and smoker status.

<br>

```{r model 3}
med_model3 <- lm(charges ~ obese*smoker + age + children, data = train)
summary(med_model3)
plot(med_model3)
```

<br>

This third model is much better than the first two, with an R-squared of .8393, implying about 84% of the variation in charges can be explained by the independent variables. Our Residual vs Fitted plot appears a little better, although the Normal Q-Q plot still indicates some problems with our fit.

##### We can interpret the model as follows:

* For an obese non-smoker, we can expect $54.04 higher charges on average
* For a non-obese smoker, we can expect $13,077.17 higher charges on average
* For each year increase in age, we can expect a $272.33 increase in average charges
* For a person who is an obese smoker, we can expect $54.04 + $13,077.17 + $19,481.61 = $32,612.82 higher charges on average

<br>

#### Finally, let's make predictions

```{r predictions}
test$predicted <- predict(med_model3, newdata = test)
test %>%
  ggplot() + geom_point(aes(predicted, charges, color = smoker, shape = obese)) +
  geom_abline(color = "red") + ggtitle("Prediction vs Real Values")
```

<br>

It looks like our model is a pretty solid fit for our test data and our results are pretty accurate.

```{r residuals}
# calculating residuals 
test$residuals <- test$charges - test$predicted

# plot residuals 
test %>%
  ggplot() + geom_pointrange(aes(predicted, residuals, ymin = 0, ymax = residuals, color = smoker, shape = obese)) +
  geom_hline(yintercept = 0) + ggtitle("Residuals vs Fitted Values")
```

<br>

Overall, the residuals look pretty good, although there are several high residuals that could be caused by factors not included in the original data set. 
